---
title: ' Class Exercise'
author: "Quinhas Fernandes"
date: "02/05/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(uwIntroStats)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(foreign)
library(haven)
knitr::opts_chunk$set(echo = TRUE, warning=TRUE, message=TRUE)

link="https://github.com/quinhasf/pubpol-599/raw/master/ape_analysis.dta"

chw_fp <- read_dta(url(link))
```


## Class Exercise: Complete adding the elements missing in the last plot - lolliplot

```{r ggplot1, eval=FALSE}
chw_pf1 <- data.frame(chw_fp[c('province', 'ape_contrib')])

tableFreq=as.data.frame(chw_pf1)

names(tableFreq)=c("province","ape_contrib")
#you have:
tableFreq = tableFreq[order(-tableFreq$ape_contrib),]

tableFreq  
```


```{r lolli_gg, eval=FALSE}
base = ggplot(tableFreq, aes(x=province,ape_contrib)) 
lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = province, 
                                   yend = ape_contrib, 
                                   xend = province), color = "grey50") 
lolliplot1 + geom_point()
```


```{r lolli_ord_gg, eval=FALSE}

tableFreqO=tableFreq[order(tableFreq$ape_contrib),]


base = ggplot(tableFreqO, aes(province,ape_contrib)) 
lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = province, 
                                   yend = ape_contrib, 
                                   xend = province), color = "gray") 
lolliplot2 = lolliplot1 + geom_point()
lolliplot2 + scale_x_discrete(limits=tableFreqO$province) 
```


```{r lolli_axis_gg, eval=FALSE}
# new variable
tableFreqO$gap=tableFreqO$ape_contrib-10

# plot the new variable
base = ggplot(tableFreqO, aes(province,gap)) 

lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = province, 
                                   yend = gap, 
                                   xend = province), color = "gray") 
lolliplot2 = lolliplot1 + geom_point()
lolliplot2 + scale_x_discrete(limits=tableFreqO$province) 
```


```{r lolli_gg_color, eval=FALSE}
# a new column for color
tableFreqO$PositiveGap=ifelse(tableFreqO$gap>0,T,F)

# add new aesthetics 'color'
base = ggplot(tableFreqO, aes(province,gap,
                              color=PositiveGap)) #change
lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = province, 
                                   yend = gap, 
                                   xend = province), color = "gray") 
lolliplot2 = lolliplot1 + geom_point()
lolliplot2 + scale_x_discrete(limits=tableFreqO$province) # key element
```


```{r nudge_gg, eval=FALSE}

tableFreqO$TargetGap=ifelse(tableFreqO$gap>0,"Below Target","Above Target")

base = ggplot(tableFreqO, aes(province,gap,color=TargetGap,
                              label = round(gap,1))) 
lolliplot1=base + geom_segment(aes(y =0, 
                                   x = province, 
                                   yend = gap, 
                                   xend = province), color = "gray") 
lolliplot2=lolliplot1 + geom_point() 

lolliplot3= lolliplot2 + scale_x_discrete(limits=tableFreqO$province) 

lollitplot4 = lolliplot3 + geom_text(nudge_x=0.3, nudge_y=0.1,show.legend = T) +
              labs(title = "Mozambique Community Health workers contribuition on Family Planning",
                   subtitle = "2017 Province level GAP on CHW contribuition",
                    x ="Province", 
                    y = "% of the FP GAP",
                    caption = "Fig.1: Represents each province GAP to achieve the 10% target (centered at 0) in 2017.
Provinces are ploted from low to high perfomance.

Source:health information system") +
            theme(panel.background = element_rect(fill = "white",
                                                    colour = "black"),
                    plot.caption = element_text(hjust = 0), 
                    plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust=0.5)) +
              geom_hline(yintercept=0,
                            linetype="solid",
                            color = "blue",
                            size= 0.9,
                            alpha= 0.8)+
  theme(legend.box.just = c("right","center")) 

lollitplot4 
```


```{r getData, eval=FALSE}
link='https://github.com/EvansDataScience/VisualAnalytics_2_tabularData/raw/master/data/eduwa.rda'

#getting the data TABLE from the file in the cloud:
load(file=url(link))
knitr::opts_chunk$set(echo = TRUE, warning=TRUE, message=TRUE)

library(uwIntroStats)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(foreign)
library(haven)
```

## Class Exercise: Make sure our box plot follows the same design approach and include all the elements as in the bar plot for nominal data

```{r ordinalTable, eval=FALSE}
table(eduwa$High.Grade,exclude = 'nothing')
```

```{r tableAsDF, eval=FALSE}
frqTabO=as.data.frame(prop.table(table(eduwa$High.Grade)))
names(frqTabO)=c('grade','pct')
frqTabO
```

```{r, eval= FALSE}
box1 = ggplot(eduwa, aes(y=as.numeric(High.Grade))) 
box1 = box1 + geom_boxplot(varwidth=T, fill="skyblue1") + coord_flip() 

ordLabels= levels(eduwa$High.Grade)

box2 = box1 + scale_y_continuous(labels=ordLabels,breaks=1:15)

box3 = box2 + 
    labs(title="2017 King County Students Grade Distribution", 
         subtitle="Box-plot of students high-grade distribution",
         caption="Source: Seattle Public Schools",
         y="Schools Grade") +
   theme(panel.background = element_rect(fill = "white",
                                                    colour = "blue"),
                    plot.caption = element_text(hjust = 0), 
                    plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust=0.5)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
box3
```

## Class Exercise: Make a histogram for the variable FREE LUNCH, and make sure it has all the right elements, and get rid of unnecessary elements.

```{r, eval=FALSE }

eduwa3=eduwa[complete.cases(eduwa$Reduced.Lunch),]

#ggplot
base= ggplot(eduwa3,aes(x = Free.Lunch))  
hist1= base + geom_histogram(binwidth = 100,boundary=0, fill="gray20") #changing width
hist2= hist1 + stat_bin(binwidth = 100, aes(label=..count..), 
                  geom = "text",boundary = 0,vjust=-0.5) +
                  labs(title="2017 Schools distribuition of Free Lunch", 
                           subtitle="Histogram of frequencies",
                       caption = "Source: Seattle Public Schools",
                       x= "Free Lunch",
                       y="Number of meals (counts)") +
                       xlim(c(0, 1800)) +
                 scale_x_continuous(breaks = c(0, 400, 800, 1200, 1600))
hist2
```

## Class Exercise: Do some research and make a histogram and a density plot using ggplot for the variable we just used above.

```{r removeMissing2, eval=FALSE}

eduwa4=eduwa[complete.cases(eduwa$Student.Teacher.Ratio),]

meanValues=tapply(eduwa4$Student.Teacher.Ratio, 
                  eduwa4$County, 
                  mean)
eduwa5 = as.data.frame(meanValues)
```

## histogram

```{r, eval=FALSE}
#ggplot
base1= ggplot(eduwa5,aes(x=meanValues)) 
histogram1= base1 + geom_histogram( binwidth = 2,boundary=0, fill="turquoise3") 
histogram2= histogram1 + stat_bin(binwidth = 2, aes(label=..count..), 
                  geom = "text",boundary = 0,vjust=-0.5) +
                  labs(title="2017 Schools students teacher ratio", 
                       caption = "Source: Seattle Public Schools",
                       x= "Mean Students Teacher ratio",
                       y="Frequency") +
                       xlim(c(10, 24)) +
          scale_x_continuous(breaks = c(10,12,14, 16, 18,20, 22, 24)) +
          geom_vline(aes(xintercept= mean(meanValues), 
                         linetype = "mean"), 
                        colour= 'red') +
          geom_vline(aes(xintercept= median(meanValues), 
                         linetype = "median"), 
                         colour= 'blue') +
          scale_linetype_manual(name = "LEGEND", values = c(2, 2), 
                       guide = guide_legend(override.aes = list(color = c("blue", "red")))) +
          theme(legend.justification =c(lty=0.5,lwd=0.5))
                
histogram2
```

## Density

```{r, eval=FALSE}

basemv= ggplot(eduwa5,aes(x=meanValues)) 

density1 = basemv + geom_density( stat="density", color="gray50") +
           theme(panel.background = element_rect(fill = "white",
                                                    colour = "black"),
                    plot.caption = element_text(hjust = 0), 
                    plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust=0.5)) +
          labs(title="2017 Schools students teacher ratio", 
                       caption = "Source: Seattle Public Schools",
                       x= "Mean Students Teacher ratio",
                       y="Frequency") +
                       xlim(c(8, 28)) + 
          geom_vline(aes(xintercept= mean(meanValues), 
                       linetype = "mean"), 
                       colour= 'red') +
          geom_vline(aes(xintercept= median(meanValues), 
                       linetype = "median"), 
                       colour= 'blue') +
          scale_linetype_manual(name = "LEGEND", values = c(12, 12), 
                      guide = guide_legend(override.aes = list(color = c("blue", "red")))) +
          theme(legend.position = c(lty=0.9,lwd=0.6))
  
density1
```
                    


